---
title: "Data Frames in R"
author: "Arie Twigt"
date: "4/7/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


# Importing the Data Frame

We will import the **German Data** data, that we will use for this project:

```{r}
# import data
german_data_clean <- read.csv("data/german_data_clean.csv")
```

> This is imported with the standard `base` functions


# Analyzing the Data Frame

## Check the columns


With the `names()` functions, we will get all the variables of the data frame.

```{r}
names(german_data_clean)
```


With the `$` operator, we can select a column in a Data Frame.

```{r eval=FALSE}
german_data_clean$credit_amount
```


Apply the `mean()` function for a column in the Data Frame:

```{r}
mean(german_data_clean$credit_amount)
```


Wrap a function over another function:

```{r}
round(mean(german_data_clean$credit_amount), 2)
```


Provide a statistical summary of a column with the `summary()` function:

```{r}
summary(german_data_clean$credit_amount)
```


Provide a summary over all columns in the Data Frame


```{r}
summary(german_data_clean)
```


Show the different values and occurances of a `character` column


```{r}
summary(factor(german_data_clean$purpose))
```

```{r}
table(german_data_clean$purpose)
```

Show the unique values of a `character` column:

```{r}
unique(german_data_clean$purpose)

```


```{r}
unique(german_data_clean$credit_amount)
```


With the `head()` function we can show the first 6 values of a vector

```{r}
head(german_data_clean$credit_amount, 10)
```


Show the last items with the `tail()` function

```{r}
tail(german_data_clean$credit_amount, 10)
```


With the `sort()` function, we can sort a vector:

```{r}
sort(german_data_clean$credit_amount)
```

With the `decreased=TRUE` parameter, we can sort on a *descending* order





With the `head()` and `sort()` function combined, we can create a **top 10**

```{r}
head(sort(german_data_clean$credit_amount, decreasing=TRUE), 10)
```

Combined `head`, `sort` and `table` functions to create a top list with names and values


```{r}
head(sort(table(german_data_clean$purpose), decreasing=TRUE), 3)
```

### Exploring the dimensions of a Table:


Show the number of rows in a Data Frame with the `nrow` function

```{r}
nrow(german_data_clean)
```


Show the number of columns in a Data Frame with the `ncol` function


```{r}
ncol(german_data_clean)
```

Show the *dimensions* of the Data Frame (both rows and columns) with the `dim` function

```{r}
dim(german_data_clean)
```


```{r}
str(german_data_clean)
```


## Explorering the data frame with visualisations


### Barplot

```{r}
barplot(c(1,2,3,4), names.arg = c("a", "b", "c", "d"))
```



```{r}
barplot(table(german_data_clean$purpose))
```
Example named vectors:


```{r}
ages <- c(20, 30, 40)
names(ages) <- c("arie", "jaap", "dirk")
sum(ages )
```


Creating a sorted barplot


```{r}
barplot(
  head(
    sort(
       table(german_data_clean$purpose), decreasing=TRUE)),
  col = "blue",  main="Top 5 Purposes", xlab="Purpose", ylab = "#")
```




```{r}

sort(table(german_data_clean$purpose), decreasing=TRUE)
```


### Boxplot


```{r}
standard_color <- "red"

boxplot(german_data_clean$credit_amount, col=standard_color, outline = FALSE, notch = TRUE, main="Credit Amount")
```



### Pie chart



```{r}
pie(table(german_data_clean$purpose))
```


```{r}
unique(german_data_clean$purpose)
custom_colours <- c("darkred", "red", "darkblue", "blue", "darkgreen", "green", "yellow", "orange", "gray", "darkgray")



```


```{r}
print(unique(german_data_clean$purpose))
print("---")
print(custom_colours)
```


```{r}
pie(table(german_data_clean$purpose), col = custom_colours)
```
### Histogram


```{r}
hist(german_data_clean$duration_months)
```


```{r}
hist(german_data_clean$duration_months, main = "Overview months", xlab = "Duration months", breaks = 5)
```


### Using a density plot in R


```{r}
plot(density(german_data_clean$credit_amount))
```



```{r}
plot(german_data_clean$credit_amount)
```


## Slicing and Dicing in R


### Selecting columnns in a Data Frame

```{r}
selected_column <- 'duration_months'

german_data_clean[selected_column]

```


```{r}
german_data_clean[1,]
```

Selecting the first 10 rows:

```{r}
german_data_clean[1:10,]
```
c()

```{r}
german_data_clean[1:10,1:3]
```




```{r}
selected_rows <- c(1:10, 200:300)
```

```{r}
german_data_clean[selected_rows,1:3]
```


```{r}
selected_columns <- c("duration_months", "purpose", "credit_amount", "response")

german_data_clean[selected_rows, selected_columns]
```


### Filtering of Data Frame



A filter applied to a vector

```{r}
number_vector <- c(3, 5, 7, 1, 8, 2, 9, 0, 3, 6)
number_vector > 5
```


```{r}
number_vector[number_vector > 5]
```

```{r}
months_filter <- german_data_clean$duration_months > 50

german_data_clean[months_filter, ]

```


```{r}
months_filter <- german_data_clean$duration_months > 50
purpose_business <- german_data_clean$purpose == "business"


german_data_clean[months_filter & purpose_business, ]

# OR
#german_data_clean[months_filter & purpose_business, ]

```


```{r}
months_filter <- german_data_clean$duration_months > 50
purpose_business <- german_data_clean$purpose %in% c("car (new)", "car (used)")


german_data_clean[months_filter & purpose_business, ]
```



## Adding calculated columns in R



```{r}
german_data_clean$amount_per_month <- german_data_clean$credit_amount / german_data_clean$duration_months
```


Verifying the calculated colum

```{r}
View(german_data_clean[,c("duration_months", "credit_amount", "amount_per_month")])
```


Calculated column based on a condition


```{r}
german_data_clean$amount_category <- ifelse(german_data_clean$credit_amount > 10000, "Large", 
                                            ifelse(german_data_clean$credit_amount > 5000, "Medium", "Low"))
```


```{r}
german_data_clean$single_male_old <- ifelse(german_data_clean$personal_status_sex == "male   : single" &
                                             german_data_clean$age_years > 68, "yes", "no")
```


Removing a column from a DataFrame


```{r}
german_data_clean$amount_per_month <- NULL
```


```{r}
cor(german_data_clean$age_years, german_data_clean$credit_amount)
```




# Exporting data in R



## Exporting a CSV file

Create a filtered Data Frame


```{r}
months_filter <- german_data_clean$duration_months > 50
purpose_business <- german_data_clean$purpose %in% c("car (new)", "car (used)")


german_data_clean_filtered <- german_data_clean[months_filter & purpose_business, ]
```


```{r}
write.table(german_data_clean_filtered, 
          file="data/export/german_data_filtered.csv",
          sep=";",
          row.names = FALSE)
```



## Export an RDS


```{r}
saveRDS(german_data_clean_filtered, "data/export/german_data_clean_filtered.rds")
```


```{r}
my_saved_rds <- readRDS("data/export/german_data_clean_filtered.rds")
```



